---
title: "PMR Tables, Age-Sex Adjusted with ICD10 Subchapters"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

```{r, include = FALSE, message = FALSE}
# Load datasets
mort65plus = read_csv("./mort17_65plus.csv")
icd10_df = read_csv("icd10_hierarchy.csv") %>% 
  select(icd10code, sub_chapter_group, sub_chapter)
```

```{r, include = FALSE}
# Create age and sex adjusted tables
# New age group
mort65plus = mort65plus %>% 
  filter(age_value < 120) %>% 
  mutate(age_grp = cut(age_value, 
                       c(65, 70, 75, 80, 85, 90, 120), 
                       right = FALSE))

age_sex = mort65plus %>% 
  select(age_grp, sex, icd10, ra2:ra20)

# Nest data by age_grp and sex
age_sex_nest = nest(age_sex, icd10:ra20)

# Create functions from previous code to apply across each dataset
# Function that filters for falls dataset
falls_death = function(data){
  data %>% 
  filter(grepl("W", icd10)) %>%
  mutate(number = str_sub(icd10, 2, 3)) %>% 
  filter(number <= 19) %>% select(-number)
}

# Function to join icd10_df to data
join_icd10 = function(data){
  full_join(data, icd10_df, by = "icd10code")
}

# Function to turn falls dataframe into a "tidy" version
tidy = function(data){
  data = data %>% 
    gather(key = ra_position, value = icd10code, ra2:ra20)
  join_icd10(data)
}

# Function to count icd10 values for overall group
count_grp = function(data){
  data = data %>% 
    gather(key = ra_position, value = icd10code, ra2:ra20) %>% 
    na.omit()
  full_join(data, icd10_df, by = "icd10code") %>% 
    group_by(sub_chapter_group) %>% 
    summarize(n = n())
}
  
# Function to count icd10 values for those with falls
count_fall = function(data){
  data %>% na.omit() %>% 
  group_by(sub_chapter_group) %>% 
  summarize(obs = n())
}

# Function to calculate PMR and PAR from observed and expected values
# Filter out any observations less than 10
pmr = function(count1, count2, data, fall){
  full_join(count1, count2, by = "sub_chapter_group") %>% 
  mutate(prev = n / dim(data)[1],
         prev_fall = obs / dim(fall)[1],
         expect = round(prev * dim(fall)[1], digits = 2),
         pmr = round(obs / expect, digits = 2),
         se_pmr = round(sqrt(obs)/expect, digits = 2),
         l95_pmr = round(pmr - 1.96 * se_pmr, digits = 2),
         u95_pmr = round(pmr + 1.96 * se_pmr, digits = 2),
         pi01 = (dim(fall)[1] - obs) / dim(fall)[1],
         pi1 = dim(fall)[1] / dim(data)[1],
         pi0 = (dim(data)[1] - dim(fall)[1]) / dim(data)[1],
         N = dim(data)[1],
         rr = (obs/n) / 
           ((dim(fall)[1] - obs) / (dim(data)[1] - n)),
         par = ((prev_fall * (rr - 1)) / (prev_fall * (pmr - 1) + 1))) %>%
  mutate(theta = 1 - par,
         var_log_theta = (1 - pi01)/(N * pi01) - 
           (pi0 + pi1 - 2 * pi01)/(N * pi0 * pi1),
         var_theta = theta^2 * var_log_theta,
         l95_par = round(100 * (par - 1.96 * sqrt(var_theta)), digits = 2),
         u95_par = round(100 * (par + 1.96 * sqrt(var_theta)), digits = 2),
         par = round(100 * par, digits = 2)) %>% 
  arrange(desc(par)) %>% 
  filter(!str_detect(sub_chapter_group, "^S")) %>%
  filter(obs > 10, !str_detect(sub_chapter_group, "^T"))
}

adjust = function(data){
  # Unnest nested data and aggregate across obs and expected values
  summary = data %>% unnest(summary)
  count = data %>% 
    unnest(count) %>% 
    group_by(sub_chapter_group) %>% 
    summarize(n = sum(n))
  
  # Calculate adjusted pmr and 95% CI 
  temp = summary %>% 
    group_by(sub_chapter_group, nfall, ntotal) %>% 
    summarize(obs = sum(obs),
              exp = sum(expect)) %>% 
    mutate(pmr = round(obs/exp, digits = 2),
           se_pmr = round(sqrt(obs)/exp, digits = 2),
           l95_pmr = round(pmr - 1.96 * se_pmr, digits = 2),
           u95_pmr = round(pmr + 1.96 * se_pmr, digits = 2)) %>% 
    filter(!str_detect(sub_chapter_group, "^S"),
           !str_detect(sub_chapter_group, "^T"),
           exp > 1, pmr > 1.5)
  
  # Calculate PAR and arrange in order
  full_join(count, temp, by = "sub_chapter_group") %>% 
    na.omit() %>% 
    mutate(prev = n/ntotal,
           par = round(100 * ((prev * (pmr - 1)) / 
                              (prev * (pmr - 1) + 1)), digits = 2)) %>% 
    arrange(desc(par))
}

# Function to tidy up the summary
clean_summary = function(data){
  icd10_df = distinct(icd10_df, sub_chapter_group, sub_chapter)
  full_join(data, icd10_df, by = "sub_chapter_group") %>% 
    filter(!is.na(pmr)) %>%
    filter(!is.na(pmr)) %>% 
    select(sub_chapter_group, obs, exp, pmr, l95_pmr, u95_pmr, par, sub_chapter) %>% 
    unite(PMR_95CI, l95_pmr, u95_pmr, sep = ", ") %>%
    mutate(PMR_95CI = paste0("(", PMR_95CI, ")")) %>% 
    rename(ICD10 = sub_chapter_group,
           Obs = obs,
           Exp = exp,
           PMR = pmr,
           PAR = par,
           SubChapter = sub_chapter)
}
```

```{r, include = FALSE, message =  FALSE}
# Apply all functions to the nested dataframes
# Create tables that summarize PMR for each age-sex group
age_sex_nest = age_sex_nest %>% 
  mutate(falls = map(.x = data, ~falls_death(.x))) %>% 
  mutate(tidy = map(.x = falls, ~tidy(.x))) %>% 
  mutate(count = map(.x = data, ~count_grp(.x)),
         count_fall = map(.x = tidy, ~count_fall(.x))) %>% 
  mutate(summary = pmap(list(count, count_fall, data, falls), pmr))
```

```{r, include = FALSE}
# Adjust for females
female = age_sex_nest %>% 
  filter(sex == "F") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Females (ntotal = `r female$ntotal[1]`, nfalls = `r female$nfall[1]`)

```{r, echo = FALSE}
female = female %>% adjust() %>% clean_summary()

female %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

```{r, include = FALSE}
# Adjust for Males
male = age_sex_nest %>% 
  filter(sex == "M") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Males (ntotal = `r male$ntotal[1]`, nfalls = `r male$nfall[1]`)

```{r, echo = FALSE}
male = male %>% adjust() %>% clean_summary()

male %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage 

```{r, include = FALSE}
# Adjust for 65-69 years
y65 = age_sex_nest %>% 
  filter(age_grp == "[65,70)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 65-69 Years (ntotal = `r y65$ntotal[1]`, nfalls = `r y65$nfall[1]`)

```{r, echo = FALSE}
y65 = y65 %>% adjust() %>% clean_summary()

y65 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```


```{r, include = FALSE}
# Adjust for 70-74 years
y70 = age_sex_nest %>% 
  filter(age_grp == "[70,75)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 70-74 Years (ntotal = `r y70$ntotal[1]`, nfalls = `r y70$nfall[1]`)
```{r, echo = FALSE}
y70 = y70 %>% adjust() %>% clean_summary()

y70 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 75-79 years
y75 = age_sex_nest %>% 
  filter(age_grp == "[75,80)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 75-79 Years (ntotal = `r y75$ntotal[1]`, nfalls = `r y75$nfall[1]`)

```{r, echo = FALSE}
y75 = y75 %>% adjust() %>% clean_summary()

y75 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```


```{r, include = FALSE}
# Adjust for 80-84 years
y80 = age_sex_nest %>% 
  filter(age_grp == "[80,85)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 80-84 Years (n = `r y80$nfall[1]`)

```{r, echo = FALSE}
y80 = y80 %>% adjust() %>% clean_summary()

y80 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 85-89 years
y85 = age_sex_nest %>% 
  filter(age_grp == "[85,90)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 85-89 Years (n = `r y85$nfall[1]`)

```{r, echo = FALSE}
y85 = y85 %>% adjust() %>% clean_summary()

y85 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```



```{r, include = FALSE}
# Adjust for 90+ years
y90 = age_sex_nest %>% 
  filter(age_grp == "[90,120)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 90+ Years (n = `r y90$nfall[1]`)

```{r, echo = FALSE}
y90 = y90 %>% adjust() %>% clean_summary()

y90 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```