---
title: "PMR Tables, Age and Sex Adjusted with Four ICD10 Digits"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

```{r, cache = TRUE, include = FALSE}
# Load datasets
mort65plus = read_csv("../mort17_65plus.csv")
icd10_df = read_csv("../icd10_hierarchy.csv")
icd10_3digit = icd10_df %>% select(icd10code, three_digit, major)
```

```{r, include = FALSE}
# Create age and sex adjusted tables
# New age group
mort65plus = mort65plus %>% 
  filter(age_value < 120) %>% 
  mutate(age_grp = cut(age_value, 
                       c(65, 70, 75, 80, 85, 90, 120), 
                       right = FALSE))

age_sex = mort65plus %>% 
  select(age_grp, sex, icd10, ra2:ra20)

# Create functions from previous code to apply across each dataset
# Function that filters for falls dataset
falls_death = function(data){
  data %>% 
  filter(grepl("W", icd10)) %>%
  mutate(number = str_sub(icd10, 2, 3)) %>% 
  filter(number <= 19) %>% select(-number)
}

# Function to turn falls dataframe into a "tidy" version
tidy = function(data, icd10_df){
  data = data %>% 
    gather(key = ra_position, value = icd10code, ra2:ra20)
}

# Function to count icd10 values for overall group
count_grp = function(data){
  data %>% 
    gather(key = ra_position, value = icd10code, ra2:ra20) %>% 
    na.omit() %>% 
    group_by(icd10code) %>% 
    summarize(n = n())
}
  
# Function to count icd10 values for those with falls
count_fall = function(data){
  data %>% 
    na.omit() %>% 
    group_by(icd10code) %>%
    summarize(obs = n())
}

# Function to count observed and expected values
pmr_count = function(count1, count2, data, fall){
  full_join(count1, count2, by = "icd10code") %>% 
    mutate(prev = n / dim(data)[1],
         prev_fall = obs / dim(fall)[1],
         expect = round(prev * dim(fall)[1], digits = 2)) %>% 
    na.omit()
}

adjust = function(data){
  # Unnest nested data and aggregate across obs and expected values
  summary = data %>% unnest(summary)
  count = data %>% 
    unnest(count) %>% 
    group_by(icd10code) %>% 
    summarize(n = sum(n))
  
  # Calculate PMR ------------------------------------------------------ 
  temp = summary %>% 
    group_by(icd10code, nfall, ntotal) %>% 
    summarize(obs = sum(obs),
              exp = sum(expect)) %>% 
    mutate(pmr = round(obs/exp, digits = 2),
           se_pmr = round(sqrt(obs)/exp, digits = 2),
           l95_pmr = round(pmr - 1.96 * se_pmr, digits = 2),
           u95_pmr = round(pmr + 1.96 * se_pmr, digits = 2)) %>% 
    filter(!str_detect(icd10code, "^S"),
           !str_detect(icd10code, "^T0"),
           !str_detect(icd10code, "^T1"),
           exp > 1, pmr > 1.5)

  full_join(count, temp, by = "icd10code") %>% 
    na.omit() %>% 
    mutate(prev = n/ntotal,
           par = round(100 * ((prev * (pmr - 1)) / 
                              (prev * (pmr - 1) + 1)), digits = 2)) %>%
    arrange(desc(par)) %>%
    filter(par >= 0.1)
}

# Function to tidy up the summary
clean_summary = function(data){
  full_join(data, icd10_df, by = "icd10code") %>%
    filter(!is.na(pmr)) %>% 
    select(icd10code, obs, exp, pmr, l95_pmr, u95_pmr, par, short_desc) %>% 
    unite(PMR_95CI, l95_pmr, u95_pmr, sep = ", ") %>%
    mutate(PMR_95CI = paste0("(", PMR_95CI, ")")) %>% 
    rename(ICD10 = icd10code,
           Obs = obs,
           Exp = exp,
           PMR = pmr,
           PAR = par)
}
```

```{r, include = FALSE}
# Nest data by age_grp and sex
age_sex_nest = nest(age_sex, icd10:ra20)

# Apply all functions to the nested dataframes
# Create tables that summarize PMR for each age-sex group
age_sex_nest = age_sex_nest %>% 
  mutate(falls = map(.x = data, ~falls_death(.x))) %>% 
  mutate(tidy = map(.x = falls, ~tidy(.x))) %>% 
  mutate(count = map(.x = data, ~count_grp(.x)),
         count_fall = map(.x = tidy, ~count_fall(.x)),
         icd10code = "icd10code") %>% 
  mutate(summary = pmap(list(count, count_fall, data, falls), pmr_count))
```

```{r, include = FALSE}
# Adjust for females
female = age_sex_nest %>% 
  filter(sex == "F") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Females (ntotal = `r female$ntotal[1]`, nfalls = `r female$nfall[1]`)

```{r, echo = FALSE}
female = female %>% select(-icd10code) %>% adjust() %>% clean_summary()

female %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage 

```{r, include = FALSE}
# Adjust for Males
male = age_sex_nest %>% 
  filter(sex == "M") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Males (ntotal = `r male$ntotal[1]`, nfalls = `r male$nfall[1]`)

```{r, echo = FALSE}
male = male %>% select(-icd10code) %>% adjust() %>% clean_summary()

male %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

### Age and Sex Adjusted

```{r, include = FALSE}
# Calculate ntotal and nfall for adjusting age and sex
age_sex_adjusted = age_sex_nest %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))

# Adjust for age and sex
age_sex_adjusted = age_sex_adjusted %>% select(-icd10code) %>% adjust() %>% clean_summary()
```

```{r, echo = FALSE}
age_sex_adjusted %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

### Male vs. Female Dot Plot

```{r, echo = FALSE}
female1 = female %>% 
  mutate(sex = "F")

male1 = male %>% 
  mutate(sex = "M")

plot1 = rbind(female1, male1)

big_diff = plot1 %>% 
  select(ICD10, PAR, sex) %>% 
  spread(sex, PAR) %>% 
  group_by(ICD10) %>% 
  mutate(Max = max(F, M),
         Min = min(F, M),
         Diff = Max / Min - 1) %>% 
  arrange(desc(Diff)) %>% 
  na.omit() 

test = big_diff %>% 
  select(-Max, -Min) %>% 
  gather(key = sex, value = PAR, F:M)
  
test %>% ggplot(aes(PAR, reorder(ICD10, PAR))) + 
  geom_line(aes(group = ICD10)) + 
  geom_point(aes(color = sex)) + 
  labs(title = "Total PAR by ICD10 and Gender",
       y = "ICD10") + 
  theme_minimal()
```

```{r, echo = FALSE}
big_diff = plot1 %>% 
  select(ICD10, PMR, sex) %>% 
  spread(sex, PMR) %>% 
  group_by(ICD10) %>% 
  mutate(Max = max(F, M),
         Min = min(F, M),
         Diff = Max / Min - 1) %>% 
  arrange(desc(Diff)) %>% 
  na.omit() 

test = big_diff %>% 
  select(-Max, -Min) %>% 
  gather(key = sex, value = PMR, F:M)
  
test %>% ggplot(aes(PMR, reorder(ICD10, PMR))) + 
  geom_line(aes(group = ICD10)) + 
  geom_point(aes(color = sex)) + 
  labs(title = "Total PMR by ICD10 and Gender",
       y = "ICD10") + 
  theme_minimal()
```



\newpage 

```{r, include = FALSE}
# Adjust for 65-69 years
y65 = age_sex_nest %>% 
  filter(age_grp == "[65,70)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 65-69 Years (ntotal = `r y65$ntotal[1]`, nfalls = `r y65$nfall[1]`)

```{r, echo = FALSE}
y65 = y65 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y65 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 70-74 years
y70 = age_sex_nest %>% 
  filter(age_grp == "[70,75)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 70-74 Years (ntotal = `r y70$ntotal[1]`, nfalls = `r y70$nfall[1]`)

```{r, echo = FALSE}
y70 = y70 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y70 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 75-79 years
y75 = age_sex_nest %>% 
  filter(age_grp == "[75,80)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 75-79 Years (ntotal = `r y75$ntotal[1]`, nfalls = `r y75$nfall[1]`)

```{r, echo = FALSE}
y75 = y75 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y75 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 80-84 years
y80 = age_sex_nest %>% 
  filter(age_grp == "[80,85)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 80-84 Years (ntotal = `r y80$ntotal[1]`, nfalls = `r y80$nfall[1]`)

```{r, echo = FALSE}
y80 = y80 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y80 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 85-89 years
y85 = age_sex_nest %>% 
  filter(age_grp == "[85,90)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 85-89 Years (ntotal = `r y85$ntotal[1]`, nfalls = `r y85$nfall[1]`)

```{r, echo = FALSE}
y85 = y85 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y85 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

\newpage

```{r, include = FALSE}
# Adjust for 90+ years
y90 = age_sex_nest %>% 
  filter(age_grp == "[90,120)") %>% 
  mutate(ntotal = map(.x = data, ~dim(.x)[1]),
         nfall = map(.x = falls, ~dim(.x)[1])) %>% 
  mutate(ntotal = Reduce("+", ntotal),
         nfall = Reduce("+", nfall))
```

### Adjusted for Ages 90+ Years (ntotal = `r y90$ntotal[1]`, nfalls = `r y90$nfall[1]`)

```{r, echo = FALSE}
y90 = y90 %>% select(-icd10code) %>% adjust() %>% clean_summary()

y90 %>% knitr::kable(booktabs = T) %>% 
    kable_styling(latex_options = c("striped", "scale_down"))
```

### PAR differences over age groups

```{r, echo = FALSE}
# Select top ICD10 by PAR
top10_1 = top_n(y65, 10, PAR)
top10_2 = top_n(y90, 10, PAR)
top10 = rbind(top10_1, top10_2) %>% 
  distinct(ICD10)

#
y65_1 = y65 %>% 
  mutate(age_grp = "65-69") %>% 
  filter(ICD10 %in% top10$ICD10)

y70_1 = y70 %>% 
  mutate(age_grp = "70-74") %>% 
  filter(ICD10 %in% top10$ICD10)

y75_1 = y75 %>% 
  mutate(age_grp = "75-79") %>% 
  filter(ICD10 %in% top10$ICD10)

y80_1 = y80 %>% 
  mutate(age_grp = "80-84") %>% 
  filter(ICD10 %in% top10$ICD10)

y85_1 = y85 %>% 
  mutate(age_grp = "85-89") %>% 
  filter(ICD10 %in% top10$ICD10)

y90_1 = y90 %>% 
  mutate(age_grp = "90+") %>% 
  filter(ICD10 %in% top10$ICD10)

plot2 = rbind(y65_1, y70_1, y75_1, y80_1, y85_1, y90_1)

plot2 %>% ggplot(aes(x = age_grp, y = PAR)) +
  geom_line(aes(group = ICD10, color = ICD10)) + 
  labs(title = "Changes in PAR over age groups",
       x = "Age Group")  +
  theme_minimal()

```


